---
title: "Prediction Assignment Writeup"
author: "Reneiro Muñoz"
date: "10/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries 

```{r message = FALSE, warning = FALSE}
library(caret)
library(ggplot2)
library(lattice)
library(e1071)

```

## Reading the data

```{r message = FALSE, warning = FALSE}
training<-read.csv("/home/reneiro/Descargas/pml-training.csv",header = TRUE,stringsAsFactors = TRUE,na.strings = "NA")
testing<-read.csv("/home/reneiro/Descargas/pml-testing.csv",header=TRUE,stringsAsFactors = TRUE,na.strings = "NA")
```


## Data cleaning

We exclude variables with NA in the data sets

```{r message = FALSE, warning = FALSE}
library(dplyr)
training <- training[, colSums(is.na(training)) == 0]
testing<-testing[,colSums(is.na(testing)) == 0]
```

We remove the variables that are not involved in the movement record

```{r message = FALSE, warning = FALSE}
trainingset<-training[,-c(1:7)]
testing<-testing[,-c(1:7)]
```

check the dimention and type of each variable for both sets
 
```{r message = FALSE, warning = FALSE}
dim(trainingset)
sapply(trainingset,class)
dim(testing)
sapply(testing,class)
```

We select numberic variables for model building.

```{r message = FALSE, warning = FALSE}
class<-trainingset$classe
trainingset<-trainingset[ ,sapply(trainingset,class)=='numeric']
testing<-testing[,sapply(testing,class)=='numeric']

trainingset<-trainingset[,colnames(testing)]
trainingset<-mutate(trainingset,classe=class)

dim(trainingset)
dim(testing)
```


## Spliting data for training and testing within training set 

We create data partition and assign training and test set for model buidling

```{r message = FALSE, warning = FALSE}
set.seed(13737)

intrain<-createDataPartition(trainingset$classe,p=0.7,list = FALSE)

trainset<-trainingset[intrain,]
testset<-trainingset[-intrain,]

```

Next, we will try three different approaches and choose the best model to run the prediction on pml-testing data

## We set up parallel processing 

```{r warning=FALSE, message = FALSE}
library(snow)
library(parallel)
c1<-makeCluster(detectCores()-1)
```
We set up the parallel processing to speed up model development


## prediction algorithm random forest 

```{r message = FALSE, warning = FALSE}
library(randomForest)
set.seed(123456)
fitControl<-trainControl(method = "cv",number = 5,allowParallel = TRUE)
modelrf<-train(classe ~., data=trainset,method="rf",trControl=fitControl)
predictrf<-predict(modelrf,testset)
confmatrixrf<-confusionMatrix(testset$classe,predictrf)
confmatrixrf

confmatrixrf$overall[1]
```
The accuracy of this model reached >99% 

## prediction algorithm gbm

```{r  message = FALSE, warning = FALSE}
library(gbm)
set.seed(123456)
fitControl<-trainControl(method = "cv",number = 5,allowParallel = TRUE)
modelgbm<-train(classe ~ .,data=trainset,method="gbm",trControl=fitControl)
predictgbm<-predict(modelgbm,testset)
confumatrixgbm<-confusionMatrix(testset$classe,predictgbm)

confumatrixgbm$overall[1]
```
The accuracy is around 95%

## prediction algorithms Decision Tree

```{r  message = FALSE, warning = FALSE }
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
modeldt<-train(classe ~ .,data=trainset,method="rpart")
fancyRpartPlot(modeldt$finalModel)
predictdt<-predict(modeldt,testset)
confumatrixdf<-confusionMatrix(predictdt,testset$classe)
confumatrixdf$overall[1]
```
The accuracy is around 43%.

## Prediction on pml-testing data

After carrying out the tests, the best model is obtained using the random forest algorithm.

Next, We estimate the out of sample error and prediction on the 20 test cases.

```{r  message = FALSE, warning = FALSE }
error<-as.numeric(1-confmatrixrf$overall[1])
error

prediction<-predict(modelrf,testing)
prediction

stopCluster(c1)
```


## Data source:
http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har
